\rule{350pt}{5pt}\\ \textbf{An Extension of Optimal Teaching For Online Perceptrons} \\ \rule{350pt}{1pt}

### 2017-05-05 paper_final.tex
Checked and cleaned up all proofs

\begin{abstract}
  Optimal teaching is the task of constructing an efficient training set
  that teaches a learner some desired concept inside its concept class. Before
  \cite{perceptron}, analysis of optimal teaching was only focused on batch
  learners. \cite{perceptron} looked at the optimal teaching strategy of an
  online Perceptron with varying degrees of knowledge of the initial state of
  the learner. In this paper, we extend these results to environments where
  there is even less knowledge of the Perceptron's initial state, but the
  teacher is given extra powers to compensate. We present an efficient algorithm
  to force the learner into a desired state with no knowledge of the learner's
  initial state. We also analyze a setting in which a teacher has the ability to
  lie.
\end{abstract}

\section{Introduction}
In the real world, there are often scenarios where a teacher wishes to
embed a desired concept within a learner. Sometimes, the teacher can only
interact with the learner by providing the learner with a set of training examples.
The task of creating the smallest possible set of examples that
produces the desired concept in the learner is known as optimal teaching
\cite{machine_teaching} \cite{teaching_dimension}\cite{td_linear}. It has applications in education and security \cite{security}\cite{education} \cite{poisoning}. \cite{perceptron} looked for
the first time at optimal teaching in an online environment by proving the
optimal teaching dimension in an online Perceptron. In this paper we look at the
optimal teaching of the same perceptron, but with a teacher that also possess
the ability to use unlabeled examples. An unlabeled example gives the teacher
knowledge of the learner's prediction about that example without altering the
learner's state. This is a reasonable power for any real-world teacher to
posses. We show that this added ability allows the teacher to teach an
online perceptron in a zero-knowledge scenario that otherwise would have been
difficult.

@@

\section{Teaching Dimension}
\cite{perceptron} used two metrics when analyzing their teaching strategies. The first was the exact 
teaching dimension: the length of the shortest training sequence that teaches a learner the desired concept.

\subsection{$TD_{exact}$}

Formally, let $\mathbb{S}$ be the space of finite training 
sequences. Then let $\mathcal{A}_{w_0}^{\eta} : \mathbb{S} \to \mathbb{R}^d$ be a function the gives the 
resultant weight vector after training on a sequence in $\mathbb{S}$, where $w_0$ is the initial weight 
vector, and $\eta$ is the learning rate. Then if the teacher which to teach a
concept $w^*$, since $w^*$ describes a linear decision boundary for any $c > 0$,
$cw^*$ is parameterizes the same concept as $w^*$. So the exact teaching dimension is 
$$
TD_{exact}(w^*, \mathcal{A}_{w_0}^\eta) = \min\{ \lvert S \rvert : 
\mathcal{A}_{w_0}^\eta(S) = cw^* \text{ for some } c >0 \}
$$

\cite{perceptron} showed that when $\eta$ and $w_0$ where known, $TD_{exact}(w^*, \mathcal{A}_{w_0}^\eta) = 1$

\subsection{$TD_\epsilon$}
\cite{perceptron} also defined an approximate teaching dimension for cases where
the exact teaching dimension was infinite, but one could approximately teach a concept within $\epsilon$ distance. For perceptrons, the distance metric used was simply the 1 minus the cosine angle between $w^*$ and $\learn(S)$
$$
1 -\frac{\langle \learn(S), w^* \rangle}{\lVert \learn(s) \rVert \lVert w^* \rVert} \leq \epsilon
$$

Then let $\mathbb{T}_\epsilon$ be the set of $\epsilon$-approximate teaching strategies for a decision boundary $w^*$. Then
$$
TD_\epsilon(w^*, \learn) = \min_{t \in 
\mathbb{T}_\epsilon} \max_{w_0 \in 
\mathbb{R}^d} \lvert t(w_0) \rvert
$$

\cite{perceptron} showed that when $w_0$ was unknown but $\eta$ was known, $TD_\epsilon(w^*, \learn) = 3$

@@

\textbf{Theorem 2.}
\textit{ If $w_0$ is known , $w_0 \leq b < \infty$ and $ 0 < \eta_{\min} \leq \eta \leq \eta_{\max} < \infty$ 
then $\mathbb{T}_{\epsilon}(w^*, \learn)  \leq 2 + \lceil \log_2 \frac{\eta_{\max} - \eta_{\min}}{2\epsilon\eta_{\max}}  \rceil$
}
\\
\\
\textit{Proof.}
Assume WLOG $\lVert w^* \rVert = 1$.

@@

Use this $(x_1, y_1)$ in \textbf{Algorithm 2} to get a bound on $\eta$ such that $\eta \in [\alpha , \beta]$ and $\beta - \alpha <= 2\epsilon \eta_{\max}$

@@

So
$$
y_2 \langle w_1, x_2 \rangle  = \bigg\langle \frac{2 b}{\epsilon (\beta + \alpha) /2} w^* + x_1 , w_0 - \eta x_1 \bigg\rangle
$$
$$
= \bigg\langle \frac{2 b}{\epsilon (\beta + \alpha) /2} w^* + x_1 , w_0  \bigg\rangle - \eta \lVert x_1 \rVert^2
$$
$$
\leq \bigg( \frac{2 b}{\epsilon (\beta + \alpha) /2} \lVert w^* \rVert + \lVert x_1 \rVert \bigg ) b - \eta \lVert x_1 \rVert^2
$$
$$
\leq \frac{4 b ^2}{\epsilon \eta_{\min}} - \eta \frac{4 b^2}{\epsilon^2 \eta_{\min}^2}
$$
$$
\leq \frac{4 b^2}{\epsilon \eta_{\min}}  -  \frac{4 b^2}{\epsilon^2 \eta_{\min}}
$$
which is less than $0$ for $\epsilon < 1$, so $(x_2 , y_2)$ will cause the perceptron to update.

@@

Then 
$$
\frac{\langle w_2, w^* \rangle}{\lVert w_2 \rVert \lVert w^* \rVert} = \frac{
\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} \lVert w^* \lVert ^2 + \langle w_0, w^* \rangle}{\lVert w_2 \rVert}
$$ 
$$
\geq \frac{\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} - b}{b +\frac{\eta 2  \lVert w_0 \rVert}{\epsilon (\beta + \alpha)/2} }
$$
$$
= 1 - \frac{2 b}{b +\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} }
$$
$$
= 1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{(\beta + \alpha)/2} } \epsilon
$$
$$
\geq  1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{\eta_{\max} - (\beta - \alpha) / 2} } \epsilon
$$
$$
\geq  1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{\eta_{\max} (1-\epsilon)} } \epsilon
$$
$$
\geq 1 - \frac{2b}{b \epsilon + \frac{ 2  b}{(1-\epsilon)} } \epsilon
$$
$$
\geq 1 - \epsilon
$$

Therefore 
$\mathbb{T}_\epsilon \leq  2 + \lceil \log_2 \frac{\eta_{\max} - \eta_{\min}}{2\epsilon\eta_{\max}}  \rceil$

\section{Teaching Dimension for unknown $\eta$ and $w_0$}
Here we present an algorithm that teaches the desired concept to the Perceptron learner when we know neither the learning rate, nor the initial weight vector of
our perceptron learner, in ? steps.

@@

Now that we have a bound around $\eta$, we get to the main result.
\\
\\
\textbf{Theorem 4.} 
\textit{With unknown $w_0$ such that $\lVert w_0 \rVert \leq b < \infty$
and unknown $\eta$ such that $0 < \eta_{\min} \leq \eta \leq \eta_{\max} < \infty$ then 
$
\mathbb{T}_\epsilon(w^* , \learn) \leq 3 + ?
$
}
\\
\\
\textit{Proof.}
First using \textbf{Algorithm 3} bound $\eta$ such that $ \alpha \leq \eta \leq \beta$ and $\beta - \alpha \leq 2\epsilon \eta_{\max}$

Then WLOG assume $\lVert w^8 \rVert = 1$. Then pick an $x_1$ such that 
$\lVert x_1 \rVert = \frac{2b}{\epsilon (\beta + \alpha)/2}$ and $\langle w^*, x_1 \rangle = 0$. Such an $x_1$ always exists. Then, let $x'_1 = -x_1.$ Then $y_1 = y'_1 = \text{sign}(\langle w*, x_1 \rangle) = -1$.  Then since $y_1 = y'_1$ then the perceptron must make an error on one of the examples. So first give $(x_1, y_1)$ to the perceptron.
\underline{Case 1}: If $(x_1, y_1)$ then $w_1 = w_0 - \eta x_1$. 
Then select at the second iteration an $x_2 = \frac{2b}{\epsilon (\beta + \alpha)/2}$. Using the same logic from theorem 2, we can see that this will give us our desired $\epsilon$ approximate $w_2$.

@@

In Case 1 we require 2 examples and and Case 2 we require 3, so including the examples needed to bound $\eta$ we get that 
$$
\mathbb{T}_\epsilon(w^*, \learn) \leq 3 + ?
$$

@@

\textbf{Case 1:} The perceptron incorrectly predicts $y = 1$ for $x_1$.

@@

Then 
$$y_2 \langle w_1, 2x_1 \rangle
$$
$$
 = \langle w_0, 2x_1 \rangle - 2\frac{(\eta b(2-\epsilon))^2}{(\eta_{\min}\epsilon)^2}
 \leq 2b\frac{\eta b(2-\epsilon)}{\eta_{\min}\epsilon}
-  2\frac{(\eta b(2-\epsilon))^2}{(\eta_{\min}\epsilon)^2} 
$$
$$
\leq 2\frac{\eta b^2(2-\epsilon)}{\eta_{\min}\epsilon}
-  2\frac{\eta (b(2-\epsilon))^2}{\eta_{\min}\epsilon^2}
< 0 \qquad \text{ for } \epsilon < 0
$$
Therefore $(x_2, y_2)$ causes and update and
$w_2 =  w_0 + \frac{\eta b(2-\epsilon)}{\eta_{\min} \epsilon}w^* $

@@

\textbf{Case 2:} $(x_1, y_1)$ does not cause an update.

@@

>>>>

\begin{abstract}
  Optimal teaching is the task of constructing an efficient training set
  that teaches a learner some desired concept inside its concept class. Before
  \cite{perceptron}, analysis of optimal teaching was only focused on batch
  learners. \cite{perceptron} looked at the optimal teaching strategy of an
  online Perceptron with varying degrees of knowledge of the initial state of
  the learner. In this paper, we extend these results to environments where
  there is even less knowledge of the Perceptron's initial state, but the
  teacher is given extra powers to compensate. We present two teaching strategies that take advantage of unlabeled examples to teach a perceptron. We also analyze a another strategy in which a teacher has the ability to
  lie.
\end{abstract}

\section{Introduction}
In the real world, there are often scenarios where a teacher wishes to
embed a desired concept within a learner. Sometimes, the teacher can only
interact with the learner by providing the learner with a set of training examples.
The task of creating the smallest possible set of examples that
produces the desired concept in the learner is known as optimal teaching
\cite{machine_teaching} \cite{teaching_dimension}\cite{td_linear}. It has applications in education and security \cite{security}\cite{education} \cite{poisoning}. \cite{perceptron} looked for
the first time at optimal teaching in an online environment by finding the teaching dimension of an online perceptron. In this paper we look at the
optimal teaching of the same perceptron, but with a teacher that also possess
the ability to use unlabeled examples. An unlabeled example gives the teacher
knowledge of the learner's prediction about that example without altering the
learner's state. This is a reasonable power for any real-world teacher to
posses. We show that this added ability, with the exclusion of some worst case scenarios, allows the teacher to teach an
online perceptron in a zero-knowledge scenario that otherwise would have been
difficult.

@@

\section{Approximate Teaching Dimension}
\cite{perceptron}  defined an approximate teaching dimension for cases where
the exact teaching dimension was infinite, but one could approximately teach a concept within $\epsilon$ distance. For perceptrons, the distance metric used was simply the 1 minus the cosine angle between $w^*$ and $\learn(S)$
$$
1 -\frac{\langle \learn(S), w^* \rangle}{\lVert \learn(s) \rVert \lVert w^* \rVert} \leq \epsilon
$$

Then let $\mathbb{T}_\epsilon$ be the set of $\epsilon$-approximate teaching strategies for a decision boundary $w^*$. Then
$$
TD_\epsilon(w^*, \learn) = \min_{t \in 
\mathbb{T}_\epsilon} \max_{w_0 \in 
\mathbb{R}^d} \lvert t(w_0) \rvert
$$
where $\learn$ is the learning algorithm and $S$ is the training sequence. 
\cite{perceptron} showed that when $w_0$ was unknown but $\eta$ was known, $TD_\epsilon(w^*, \learn) = 3$

@@

\textbf{Theorem 2.}
\textit{ If $w_0$ is known , $\lVert w_0 \rVert \leq b < \infty$ and $ 0 < \eta_{\min} \leq \eta \leq \eta_{\max} < \infty$ 
then $\mathbb{T}_{\epsilon}(w^*, \learn)  \leq 2 + \lceil \log_2 \frac{\eta_{\max} - \eta_{\min}}{2\epsilon\eta_{\max}}  \rceil$
}
\\
\\
\textit{Proof.}
Assume WLOG $\lVert w^* \rVert = 1$.

@@

Use this $(x_1, y_1)$ in \textbf{Algorithm 2} to get a bound on $\eta$ such that $\eta \in [\alpha , \beta]$ and $\beta - \alpha \leq 2\epsilon \eta_{\max}$

@@

So
$$
y_2 \langle w_1, x_2 \rangle  = \bigg\langle \frac{2 b}{\epsilon (\beta + \alpha) /2} w^* + x_1 , w_0 - \eta x_1 \bigg\rangle
= \bigg\langle \frac{2 b}{\epsilon (\beta + \alpha) /2} w^* + x_1 , w_0  \bigg\rangle - \eta \lVert x_1 \rVert^2
$$
$$
\leq \bigg( \frac{2 b}{\epsilon (\beta + \alpha) /2} \lVert w^* \rVert + \lVert x_1 \rVert \bigg ) b - \eta \lVert x_1 \rVert^2
\leq \frac{4 b ^2}{\epsilon \eta_{\min}} - \eta \frac{4 b^2}{\epsilon^2 \eta_{\min}^2}
\leq \frac{4 b^2}{\epsilon \eta_{\min}}  -  \frac{4 b^2}{\epsilon^2 \eta_{\min}}
$$
which is less than $0$ for $\epsilon < 1$, so $(x_2 , y_2)$ will cause the perceptron to update.

@@

Then 
$$
\frac{\langle w_2, w^* \rangle}{\lVert w_2 \rVert \lVert w^* \rVert} = \frac{
\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} \lVert w^* \lVert ^2 + \langle w_0, w^* \rangle}{\lVert w_2 \rVert}
$$ 
$$
\geq \frac{\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} - b}{b +\frac{\eta 2  \lVert w_0 \rVert}{\epsilon (\beta + \alpha)/2} }
= 1 - \frac{2 b}{b +\frac{\eta 2  b}{\epsilon (\beta + \alpha)/2} }
= 1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{(\beta + \alpha)/2} } \epsilon
$$
$$
\geq  1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{\eta_{\max} - (\beta - \alpha) / 2} } \epsilon
\geq  1 - \frac{2b}{b \epsilon + \frac{\eta 2  b}{\eta_{\max} (1-\epsilon)} } \epsilon
\geq 1 - \frac{2b}{b \epsilon + \frac{ 2  b}{(1-\epsilon)} } \epsilon
$$
$$
\geq 1 - \epsilon
$$

Therefore 
$\mathbb{T}_\epsilon(w^* ,\learn) \leq  2 + \lceil \log_2 \frac{\eta_{\max} - \eta_{\min}}{2\epsilon\eta_{\max}}  \rceil$

\subsection{Tightness of Bound}
We believe the bound in \textbf{Theorem 2} is tight. There are examples for which a single labeled example will not be able to guarantee our $\epsilon$ upper bound without a better bound on the learning rate. And we believe \textbf{Algorithm 2} is optimal for bounding the learning rate. Further work will need to be done to prove the tightness of \textbf{Theorem 2}.

\section{Teaching Dimension for unknown $\eta$ and $w_0$}
Here we present an algorithm that teaches the desired concept to the Perceptron learner when we know neither the learning rate, nor the initial weight vector of
our perceptron learner.

@@

Now that we have a bound around $\eta$, we get to the main result.
\\
\\
\textbf{Theorem 4.} 
\textit{With unknown $w_0$ such that $\lVert w_0 \rVert \leq b < \infty$
and unknown $\eta$ such that $0 < \eta_{\min} \leq \eta \leq \eta_{\max} < \infty$ then 
$
\mathbb{T}_\epsilon(w^* , \learn) \leq 3 + T_{\text{bound}}(2\epsilon \eta_{\max})
$
where $T_{\text{bound}}(2\epsilon \eta_{\max})$ is the worst case number of examples needed to bound $\eta \in [\alpha , \beta]$ s.t. $\beta - \alpha \leq 2\epsilon \eta_{\max}$
}
\\
\\
\textit{Proof.}
First bound $\eta$ such that $ \alpha \leq \eta \leq \beta$ and $\beta - \alpha \leq 2\epsilon \eta_{\max}$

Then WLOG assume $\lVert w^* \rVert = 1$. Then pick an $x_1$ such that 
$\lVert x_1 \rVert = \frac{2b}{\epsilon (\beta + \alpha)/2}$ and $\langle w^*, x_1 \rangle = 0$. Such an $x_1$ always exists. Then, let $x'_1 = -x_1.$ Then $y_1 = y'_1 = \text{sign}(\langle w*, x_1 \rangle) = -1$.  Then since $y_1 = y'_1$ then the perceptron must make an error on one of the examples. So first give $(x_1, y_1)$ to the perceptron.

\underline{Case 1}: If $(x_1, y_1)$ causes an update then $w_1 = w_0 - \eta x_1$. 
Then select at the second iteration an $x_2 = \frac{2b}{\epsilon (\beta + \alpha)/2}$. Using the same logic from theorem 2, we can see that this will give us our desired $\epsilon$ approximate $w_2$.

@@

In Case 1 we require 2 examples and and Case 2 we require 3, so including the examples needed to bound $\eta$ we get that 
$$
\mathbb{T}_\epsilon(w^*, \learn) \leq 3 + T_{\text{bound}}(2\epsilon \eta_{\max})
$$

@@

\underline{Case 1:} The perceptron incorrectly predicts $y = 1$ for $x_1$.

@@

Then 
$$y_2 \langle w_1, x_2 \rangle
 = \langle w_0, 2x_1 \rangle - 2\frac{(\eta b(2-\epsilon))^2}{(\eta_{\min}\epsilon)^2}
$$
$$
 \leq 2b\frac{\eta b(2-\epsilon)}{\eta_{\min}\epsilon}
-  2\frac{(\eta b(2-\epsilon))^2}{(\eta_{\min}\epsilon)^2} 
$$
$$
\leq 2\frac{\eta b^2(2-\epsilon)}{\eta_{\min}\epsilon}
-  2\frac{\eta (b(2-\epsilon))^2}{\eta_{\min}\epsilon^2}
< 0 \qquad \text{ for } \epsilon < 0
$$
Therefore $(x_2, y_2)$ causes and update and
$w_2 =  w_0 + \frac{\eta b(2-\epsilon)}{\eta_{\min} \epsilon}w^* $

@@

\underline{Case 2:} $(x_1, y_1)$ does not cause an update.

@@

\section{Conclusion}
